---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
# packages
library(tswge)
library(vars)
library(RColorBrewer) #for "fan" forecasts
library(GGally)
library(quantmod)
library(nnfor)
library(dplyr)
library(lubridate)

# if there is an issue with TSWGE - use this
### gitHub TSWGE
# library(devtools)
# install_github("BivinSadler/tswge")
# # WHEN PROMPTED FOR UPDATES = SELECT 3 = NONE
# # make sure to terminate R and restart session
# library(tswge)
# # exists('roll.win.rmse.wge') 
### If this returns = TRUE then you have the GitHub version of TSWGE, and need the long-hand ASE function call 


# If it is false = then you are using the current TSWGE package  from:
# # remove.packages('tswge')
# # install.packages('tswge')

```


```{r}
set.seed(17)
setwd("~/00 SMU/07 time series/project/datasources")
walSales = read.csv("walmart_sales.csv",header=T,stringsAsFactors = TRUE)
inDates = read.csv("dates_in.csv",header=T)
walStock = read.csv("walmart_stock.csv",header=T,stringsAsFactors = TRUE)
```


```{r}
inDates2 = inDates %>%
  select(FISCAL_WEEK_END
         ,DATE_ID
         ,FISCAL_YEAR_ID
         ,FISCAL_YEAR_START)
inDates = inDates2

inDates$Date = inDates$DATE_ID
length(walSales$Weekly_Sales)
```

```{r}
stock_df = merge(x=walStock,y=inDates,all.x=TRUE)
head(stock_df)
```

```{r}
stock_df$Date = as.Date(mdy(stock_df$Date))

stock_weekly_stg = data.frame(stock_df %>%
                                group_by (FISCAL_WEEK_END) %>%
                                summarize(
                                  avgOpen = mean(Open, na.rm=TRUE),
                                  avgHigh = mean(High, na.rm=TRUE),
                                  avgLow = mean(Low, na.rm=TRUE),
                                  avgClose = mean(Close, na.rm=TRUE),
                                  avgVolume = mean(Volume, na.rm=TRUE),
                                  FISCAL_YEAR_ID = max(FISCAL_YEAR_ID)
                                ))
head(stock_weekly_stg)

stock_weekly = stock_weekly_stg

stock_weekly$FISCAL_WEEK_END = as.factor(stock_weekly$FISCAL_WEEK_END)
stock_weekly$FISCAL_WEEK_END = strptime(stock_weekly$FISCAL_WEEK_END, format="%m/%d/%Y")
stock_weekly$FISCAL_WEEK_END = as.Date(stock_weekly$FISCAL_WEEK_END, format="%m/%d/%Y")
str(stock_weekly)

stock_weekly = stock_weekly[order(as.Date(stock_weekly$FISCAL_WEEK_END)),]
stock_weekly
```


```{r}
# stock_weekly$Date = as.Date(mdy(stock_weekly$Date))
head(stock_weekly)

stock_weekly %>% group_by(FISCAL_YEAR_ID) %>% summarize(count=n())

stock_weekly = stock_weekly[which(stock_weekly$FISCAL_YEAR_ID == '2,010'|
                                    stock_weekly$FISCAL_YEAR_ID == '2,011'|
                                    stock_weekly$FISCAL_YEAR_ID == '2,012'), ]
stock_weekly %>% group_by(FISCAL_YEAR_ID) %>% summarize(count=n())
head(stock_weekly, n=50)
stock_weekly

# stock_weekly$FISCAL_WEEK_END = as.Date(mdy(stock_df$FISCAL_WEEK_END))

```


### UNIVARIATE ANALYSIS OF STOCK DATA


## OBSERVATIONS
Obs 01. The realization shows a wandering behavior.

Obs 02. Sample autocorrelations plot is damping and there is a peak at 0 in Frequency plot.


### Step 01. plot the data
# confirm data are not white noise
```{r}
aic5.wge(stock_weekly$avgClose)
plotts.wge(stock_weekly$avgClose)
plotts.sample.wge(stock_weekly$avgClose)
```

# peaks at 0

```{r}
pacf(stock_weekly$avgClose)

priceClose = parzen.wge(stock_weekly$avgClose)

# priceClose$freq[findPeaks(priceClose$pzgram)]
```

TRAIN / TEST SPLIT DATA
```{r}
# create train, test data set (future note: would want to add validate set)
# length(stock_weekly$avgClose)
stock_train = stock_weekly[1:142,]
stock_test = stock_weekly[143:157,]
```

### MODEL 01. UNIVARIATE AR(1)
# Model ID (also, confirm the data are not white noise)
```{r}
aic5.wge(stock_train$avgClose)
```

The above code provides evidence that the data are not consistent with simply white noise.
White noise would return p = 0, and q = 0.

AIC scores suggest a correlation structure of p=1, q=0, an AR(1) model.

Next, we estimate the model parameters using AIC selected model (ML Estimates).
```{r}
mod01est = est.arma.wge(stock_train$avgClose,p=1,q=0)

mean(stock_train$avgClose)
mod01est$phi
mod01est$theta
mod01est$avar
```
# Equation for Model 01
>(1 - 0.994B) (X - 56.50996) = a(t)


# white noise evaluation of residuals
## Check 01. Visual

A visual check of the residuals' autocorrelation shows some evidence that the variance may not be consistent with white noise. (We would expect to see about 1 in 20 outside the lines in the plot.)
Noting this, we will use the Ljung-Box test to check the residuals for white noise.

Desired result = residuals consistent with white noise. We want to be sure that the noise between points in the model is not correlated. If we still have correlation, we will difference the data again.

```{r}
plotts.wge(mod01est$res)
acf(mod01est$res)
```

## Check 02. Ljung Box test
```{r}
ljung.wge(mod01est$res, p=1, q=0)
ljung.wge(mod01est$res, p=1, q=0, K=48)
```
Ljung-Box Test for both K=24 and 48 provide evidence that the residuals are consistent with noise.
We reject the null hypothesis for both, p-value > 0.05.

# forecast last 5 values using estimated parameters from above
```{r}
mod01preds = fore.arma.wge(stock_train$avgClose,phi = mod01est$phi,
                           theta = mod01est$theta,n.ahead = 5, lastn=FALSE)

ase01ST = mean((mod01preds$f - stock_test$avgClose[1:5])^2)
ase01ST
```

```{r}
# # plot forecast preds against actuals
plot(stock_test$avgClose[1:5], type='l',main = 'Model 1 Short Term Preds vs Actuals',
     ylim = c(70,80))
lines(seq(1,5), mod01preds$f, col = 'blue')

# plot(stock_weekly$avgClose[1:147], type='l',main = 'Model 1 Short Term Preds vs Actuals')
# lines(seq(143,147), mod01preds$f, col = 'blue')
```


```{r}
plot(stock_weekly$avgClose[1:147], type='l',main = 'Model 1 Short Term Preds vs Actuals',
     ylim = c(40, 80))
lines(seq(143,147), mod01preds$f, col = 'blue')
```


### MODEL 02. ARIMA(1,2,1)

We notice that the parameter estimate for our AR(1) model has an absolute reciprocal value very close to 1, which is very nearly nonstationary. 

Difference the data (Box-Jenkins method to help with unit roots. High-pass filter.)
One way to address nonstaionary behavior due to a root on the unit circle is to difference the data until they appear stationary.
```{r}
# take first difference of the data
mod02df = artrans.wge(stock_train$avgClose, 1)
plotts.sample.wge(mod02df)
aic5.wge(mod02df)
```
After taking the first difference of the data, still shows us AR(1).


Difference the data again.
```{r}
# pass in the differenced data
mod02df1 = artrans.wge(mod02df, 1)
plotts.sample.wge(mod02df1,arlimit=TRUE)
aic5.wge(mod02df1)

```


```{r}
mod02est = est.arma.wge(mod02df1,p=1, q = 1)

mean(mod02df1)
mod02est$phi
mod02est$theta
mod02est$avar
```
# Note: if mean is positive, subtract from final model estimation.

# Equation for Model 02. ARIMA(1,2,1)
> (1 - 0.1476B)((1 - B)^2)(X(t) - 0.006225) = (1 - B)a(t)

When we originally ran this, our MA term was 0.999B.
A second pass suggests that maybe we should address the unit root in the MA component. 

Options for additional analysis could be to test another AR(p) model with additional differencing, layering in seasonality, or applying a moving average filter to the data. 


# White noise evaluation: Model 02.
## Check 01. Visual
```{r}
plotts.wge(mod02est$res)
acf(mod02est$res)
```

```{r}
ljung.wge(mod01est$res, p=1, q=1) # K=24, p-value > 0.05, fail to reject whitenoise
ljung.wge(mod01est$res, p=1, q=1, K=48) # K=24, p-value > 0.05
```
We fail to reject for both tests = good to go


# SHORT TERM FORECAST 
Predictions for last 5 values using estimated parameters from above.
```{r}
mod02preds = fore.aruma.wge(stock_train$avgClose,phi = mod02est$phi, d=2,
                           theta = mod02est$theta,n.ahead = 5, lastn=FALSE)
```

```{r}
ase02ST = mean((mod02preds$f - stock_test$avgClose[1:5])^2)
ase02ST
```

```{r}
# # plot forecast preds against actuals
plot(stock_test$avgClose[1:5], type='l',main = 'Model 2 Short Term Preds vs Actuals',
     ylim = c(70,80))
lines(seq(1,5), mod02preds$f, col = 'blue')

```

```{r}
plot(stock_weekly$avgClose[1:147], type='l',main = 'Model 2 Short Term Preds vs Actuals',
     ylim = c(40, 80))
lines(seq(143,147), mod02preds$f, col = 'blue')
```

LONG TERM FORECASTS

Model 1: AR(1): Long term forecasts
```{r}
# Long term forecast Model 1
mod01predsLT = fore.arma.wge(stock_train$avgClose,phi = mod01est$phi,
                           theta = mod01est$theta, n.ahead = 15, limits = T, lastn=FALSE)

```

Model 1: AR(1): Long term forecasts: ASE 
```{r}
ase01LT = mean((mod01predsLT$f - stock_weekly$avgClose[143:157])^2)
ase01LT
```

Model 1: AR(1): Long term forecasts: Predictions vs actuals
```{r}
# # plot forecast preds against actuals
plot(stock_test$avgClose[1:15], type='l',main = 'Model 1 Long Term Preds vs Actuals',
     ylim = c(65,78))
lines(seq(1,15), mod01predsLT$f, col = 'blue')

```
Model 1: AR(1): Long term forecasts: Predictions vs actuals
```{r}
# # plot forecast preds against actuals
plot(stock_weekly$avgClose, type='l',main = 'Model 1 LT Preds v Actuals')
lines(seq(143,157), mod01predsLT$f, col = 'blue')
```

Model 2. ARIMA(1,2,1): Long term forecasts
```{r}
# Long term forecast Model 2
mod02predsLT = fore.aruma.wge(stock_train$avgClose, phi = mod02est$phi, d = 2,
                           theta = mod02est$theta, n.ahead = 15, limits = T, lastn=FALSE)

```

Model 2. ARIMA(1,2,1): Long term forecasts: ASE
```{r}
ase02LT = mean((mod02predsLT$f - stock_weekly$avgClose[143:157])^2)
ase02LT
```

Model 2. ARIMA(1,2,1): Long term forecasts: Predictions vs actuals
```{r}
# # plot forecast preds against actuals
plot(stock_weekly$avgClose, type='l',main = 'Model 2 LT Preds v Actuals',
      ylim = c(45,85))
lines(seq(143,157), mod02predsLT$f, col = 'blue')
```

Model 2 has very poor performance. We are forecasting the trend to continue, when it very clearly does not. 
Further assessment of the data or adding aditional features may be helpful for our model.



### ROLLING WINDOWS
ROLLING WINDOW ASE FUNCTION 
```{r}
roll.win.ase.wge = function(series, horizon = 1, s = 0, d = 0, phis = 0, thetas = 0)
{

  trainingSize = length(phis) + length(thetas) + s + d + 1
  numwindows = length(series)-(trainingSize + horizon) + 1
  ASEHolder = numeric(numwindows)

  print(paste("Please Hold For a Moment, TSWGE is processing the Rolling Window ASE with", numwindows, "windows."))

  for( i in 1:numwindows)
  {

    invisible(capture.output(forecasts <- fore.aruma.wge(series[i:(i+(trainingSize-1))],phi = phis, theta = thetas, s = s, d = d,n.ahead = horizon)))

    ASE = mean((series[(trainingSize+i):(trainingSize+ i + (horizon) - 1)] - forecasts$f)^2)

    ASEHolder[i] = ASE

  }

  ASEHolder
  hist(ASEHolder, main = "ASEs for Individual Windows")
  WindowedASE = mean(ASEHolder)

  print("The Summary Statistics for the Rolling Window ASE Are:")
  print(summary(ASEHolder))
  print(paste("The Rolling Window ASE is: ",round(WindowedASE,3)))
  return(list(rwASE = WindowedASE, numwindows = numwindows, horizon = horizon, s = s, d = d, phis = phis, thetas = thetas))
}

```

Model 01: AR(1): Short term forecasts: Rolling window ASE
```{r}
# function [roll.win.rmse.wge] does not exist with "regular" TSWGE (need the gitHub version for that)
# this is the ROLLING WINDOW ASE (WITH HORIZON = 5)
roll_mod01ST_ase = roll.win.ase.wge(stock_train$avgClose, horizon = 5, phis = mod01est$phi,
                           thetas = mod01est$theta)

roll_mod01ST_ase$rwASE
sqrt(roll_mod01ST_ase$rwASE)
```


Model 01: AR(1): Short term forecasts: Rolling window ASE
```{r}
roll_mod01ST_aseRW = round(roll_mod01ST_ase$rwASE, 4)
roll_mod01ST_sqr_aseRW = round(sqrt(roll_mod01ST_ase$rwASE),4)
print(paste0('The rolling window ASE for Model 01. AR(1), short term (train df) is: ', roll_mod01ST_aseRW))
print(paste0('The square root of the rolling window ASE for Model 01. AR(1), short term (train df) is: ', roll_mod01ST_sqr_aseRW))
```


Model 01 AR(1): Short term forecasts: Rolling window RMSE
```{r}
roll_mod01ST = roll.win.rmse.wge(stock_train$avgClose, horizon = 5, phi = mod01est$phi,
                                  theta = mod01est$theta)
```


Model 01: AR(1): Short term forecasts: Rolling window RMSE
```{r}
roll_mod01ST_rmse = round(roll_mod01ST$rwRMSE,4)
print(paste0('The rolling window RMSE for Model 01. AR(1), short term is: ', roll_mod01ST_rmse))
```

Model 01: AR(1): Long term forecasts: Rolling window RMSE
```{r}
roll_mod01LT = roll.win.rmse.wge(stock_train$avgClose, horizon = 15, phi = mod01est$phi,
                                  theta = mod01est$theta)

# roll_mod01LT_rmse = round(sqrt(roll_mod01LT$rwRMSE),4)
# print(paste0('The rolling window RMSE for Model 01. AR(1), long term is: ', roll_mod01LT_rmse))

```


Model 01 AR(1): Long term forecasts: Rolling window RMSE
```{r}
roll_mod01LT_rmse = round(roll_mod01LT$rwRMSE,4)
print(paste0('The rolling window RMSE for Model 01. AR(1), long term is: ', roll_mod01LT_rmse))
```


Model 02: ARIMA(1,2,1): Short term forecasts: Rolling window RMSE
```{r}
roll_mod02ST = roll.win.rmse.wge(stock_train$avgClose, horizon = 5, phi = mod02est$phi,
                                  theta = mod02est$theta)

```
Model 02: ARIMA(1,2,1): Short term forecasts: Rolling window RMSE
```{r}
roll_mod02STrmse = round(roll_mod02ST$rwRMSE, 4)
print(paste0('The rolling window RMSE for Model 02. ARIMA(1,2,1), short term is: ', roll_mod02STrmse))
```
Model 02: ARIMA(1,2,1): Short term forecasts: Rolling window RMSE
```{r}
roll_mod02ST = roll.win.rmse.wge(stock_train$avgClose, horizon = 5, phi = mod02est$phi,
                                  theta = mod02est$theta)

```


Model 02: ARIMA(1,2,1): Short term forecasts: Rolling window RMSE
```{r}
roll_mod02ST_rmse = round(roll_mod02ST$rwRMSE,4)
print(paste0('The rolling window RMSE for Model 02. ARIMA(1,2,1), short term is: ', roll_mod02ST_rmse))

```

Model 02: ARIMA(1,2,1): Long term forecasts: Rolling window RMSE
```{r}
roll_mod02LT = roll.win.rmse.wge(stock_train$avgClose, horizon = 15, phi = mod02est$phi,
                                  theta = mod02est$theta)
```


Model 02: ARIMA(1,2,1): Long term forecasts: Rolling window RMSE
```{r}
roll_mod01LT_rmse = round(roll_mod01ST$rwRMSE,4)
print(paste0('The rolling window RMSE for Model 02. ARIMA(1,2,1), long term is: ', roll_mod01LT_rmse))

```


RMSE Comparison of forecasts (on stock data, univariate analysis).
```{r}
print(paste0('The rolling window RMSE for Model 01. AR(1), short term is: ', roll_mod01ST_rmse))
print(paste0('The rolling window RMSE for Model 01. AR(1), long term is: ', roll_mod01LT_rmse))
print(paste0('The rolling window RMSE for Model 02. ARIMA(1,2,1), short term is: ', roll_mod02ST_rmse))
print(paste0('The rolling window RMSE for Model 02. ARIMA(1,2,1), long term is: ', roll_mod01LT_rmse))

```
The AR(1) model does better both in the short term, and in the long term, when it comes to forecasting the average weekly close price of Walmart stock. 


SALES DATAFRAME CLEAN/EDA 
```{r}
# first want to prep data
# LOAD IN / PREP SALES DF
sales_df = merge(x=walSales,y=inDates,all.x=TRUE)
sales_df$Date = as.Date(mdy(sales_df$Date))
str(sales_df)
head(sales_df)
```

EDA: Dates
```{r}
# set date formatting 
sales_df$FISCAL_WEEK_END = as.factor(sales_df$FISCAL_WEEK_END)
sales_df$FISCAL_WEEK_END = strptime(sales_df$FISCAL_WEEK_END, format="%m/%d/%Y")
sales_df$FISCAL_WEEK_END = as.Date(sales_df$FISCAL_WEEK_END, format="%m/%d/%Y")
str(sales_df)

# order by FISCAL_WEEK_END (important for time series!)
sales_df = sales_df[order(as.Date(sales_df$FISCAL_WEEK_END)),]
sales_df
```

```{r}
print(length(unique(sales_df$Store)))
```

We have 45 unique stores in our dataset. 

Next, we will plot out some of the realizations to see if we have consistency across the stores.
```{r}
store1 = sales_df[which(sales_df$Store == '1'),]
plotts.sample.wge(store1$Weekly_Sales)

store1 = sales_df[which(sales_df$Store == '2'),]
plotts.sample.wge(store1$Weekly_Sales)

store1 = sales_df[which(sales_df$Store == '11'),]
plotts.sample.wge(store1$Weekly_Sales)

# Store 33 = note: this one looks a little different from the rest
store1 = sales_df[which(sales_df$Store == '33'),]
plotts.sample.wge(store1$Weekly_Sales)

store1 = sales_df[which(sales_df$Store == '45'),]
plotts.sample.wge(store1$Weekly_Sales)
```


```{r}
sales_df2 = subset(sales_df, select = -c(Date2, FISCAL_YEAR_START))
sales_df = sales_df2

colnames(sales_df)
```

Let's grab average values across all stores by week. 
```{r}
sales_df_stg = data.frame(sales_df %>%
                                group_by (FISCAL_WEEK_END) %>%
                                summarize(
                                  avgSales = mean(Weekly_Sales, na.rm=TRUE),
                                  avgTemp = mean(Temperature, na.rm=TRUE),
                                  avgFuel = mean(Fuel_Price, na.rm=TRUE),
                                  avgCPI = mean(CPI, na.rm=TRUE),
                                  avgUnemp = mean(Unemployment, na.rm=TRUE),
                                  FISCAL_YEAR_ID = max(FISCAL_YEAR_ID),
                                  Holiday_Flag = max(Holiday_Flag)
                                ))
sales_df = sales_df_stg
head(sales_df)
```

MULTIVARIATE ANALYSIS

Merge sales_df (sales info and other variables) with stock price DF.
```{r}
length(stock_weekly$avgClose) # to confirm row counts # 157
length(sales_df$avgTemp) # 124

sales_df %>% group_by(FISCAL_YEAR_ID) %>% summarize(count=n())
stock_weekly %>% group_by(FISCAL_YEAR_ID) %>% summarize(count=n())

# finding: we have more records in the stock data than in the sales data
# we will reduce the stock data to align with the sales data

mult_df = merge(x = stock_weekly, y = sales_df, by='FISCAL_WEEK_END')
                 # all.x = TRUE )
length(mult_df$avgClose) # to confirm row counts

head(mult_df)
# mult_df$FISCAL_YEAR_ID = mult_df$FISCAL_YEAR_ID.x
sales_df %>% group_by(FISCAL_YEAR_ID) %>% summarize(count=n()) # looks good = matches above
```

Plot explanatory variables.
```{r}
# colnames(mult_df)
# plotts.sample.wge(ts(mult_df11$Weekly_Sales))

# Weekly Sales
plotts.sample.wge(mult_df$avgSales) 

# Temperature
plotts.sample.wge(mult_df$avgTemp)

# Fuel Price
plotts.sample.wge(mult_df$avgFuel)

# CPI
plotts.sample.wge(mult_df$avgCPI)

# Unemployment
plotts.sample.wge(mult_df$avgUnemp)

```


### CROSS CORRELATIONS PLOTS
```{r}
# Weekly Sales
ccf(mult_df$avgClose, mult_df$avgSales, lag.max=17,
    main="Cross-correlation of Average Close and Sales")

# Temperature
ccf(mult_df$avgClose, mult_df$avgTemp, lag.max=17,
    main="Cross-correlation of Average Close and Temperature")

# Fuel Price
ccf(mult_df$avgClose, mult_df$avgFuel, lag.max=17,
    main="Cross-correlation of Average Close and Fuel Price")

# CPI
ccf(mult_df$avgClose, mult_df$avgCPI, lag.max=17,
    main="Cross-correlation of Average Close and CPI")

# Unemployment
ccf(mult_df$avgClose, mult_df$avgUnemp, lag.max=17,
    main="Cross-correlation of Average Close and Unemployment")

```

Note: Evaluation up to this point was performed on the full data, initial checks of the test dataset confirm that the data behavie in similar fashion. 

TRAIN / TEST SPLIT On new dataframe (sales and stock data) 
```{R}
# FILTER TRAINING SET = REMOVE LAST 15 OBSERVATIONS
# length(mult_df$avgClose) # 124 = remove last 15

mult_df_train = mult_df[1:109,]
mult_df_test = mult_df[110:124,]

# confirm counts
length(mult_df_train$avgClose)
length(mult_df_test$avgClose)
```


### UNIVARIATE ANALYSIS OF AVG SALES
Plot the data
```{r}
# Weekly Average Sales plots
plotts.sample.wge(mult_df_train$avgSales) 

```

Model id: Weekly average sales
```{r}
p1 = suppressWarnings(aic5.wge(mult_df_train$avgSales, p = 0:9, q = 0:9))
p1
# AIC picks p = 0, q = 3 // p = 7, q = 7
```

Model estimates [AIC values #1 = MA(3)]: Weekly average close
```{r}
sales1.est = est.arma.wge(mult_df_train$avgSales, p = 0, q = 3)
```

Even though the AIC picked an MA(3) as the model with the lowest AIC score, we know that MA models are not as useful on their own. The MA components are more useful when used alongside the AR pieces. 
Given this, we will move forward to get the model estimates for the ARMA(7,7) model = choice #2.

Model estimates [AIC values #2 = ARMA(7,7)]: Weekly average close
```{r}
sales2.est = est.arma.wge(mult_df_train$avgSales, p = 7, q = 7)
# shows evidence of unit root (1-B) term 
```

Difference data to remove the unit root.
```{r}
sales_dif1 = artrans.wge(mult_df_train$avgSales, phi.tr = 1)
```

Model id: Weekly average sales (d = 1)
```{r}
sales_dif1 = ts(sales_dif1)
p_dif1 = suppressWarnings(aic5.wge(sales_dif1, p = 0:9, q = 0:9))
p_dif1
# AIC picks AR(6) // p = 6
```

Model estimates [AIC values #1 = AR(6)]: Weekly average close with d = 1
```{r}
sales_dif1_est = est.ar.wge(sales_dif1, p = 6)
```

Forecast: Weekly average sales 
```{r}
sales_dif1_preds = fore.arma.wge(sales_dif1, phi = sales_dif1_est$phi, n.ahead = 15, lastn = FALSE, limits = F)

```
How did our model do? 
```{r}
ase_dif1 = mean((sales_dif1_preds$f - mult_df_test$avgSales[1:15])^2)
ase_dif1

# pretty good with modeling the sales data
```


### VAR MODEL
Using weekly average close and weekly average sales.
```{r}
# bind variables together 
X = cbind(mult_df_train$avgSales, mult_df_train$avgClose)
colnames(X) = c('Sales','Close')

X = na.omit(X)

# fit model
VAR_SM = VAR(X,lag.max = 17, type = 'both')

# predictions shor term 
predsST = predict(VAR_SM,n.ahead = 5)
```

```{r}
# # plot forecast preds against actuals
plot(stock_weekly$avgClose[1:147], type='l',main = 'VAR Model ST Preds v Actuals',
      ylim = c(45,85))
lines(seq(143,147), predsST$fcst$Close[,1], col = 'blue')
```



How does our model do predicting the short term?
```{r}
# preds$fcst$Sales ## useful reference - keep in 
ase_varST = mean((predsST$fcst$Close[1:5] - mult_df_test$avgClose[1:5])^2)
ase_varST
# rmse_varST = sqrt(ase_varST)
# rmse_varST
print(paste0('The ASE for the VAR model, short term is: ', round(ase_varST,4) ))
print(paste0('The square root of the ASE for the VAR model, short term is: ', round(sqrt(ase_varST),4) ))
```


Long term predictions for VAR model.
```{r}

# predictions long term 
preds = predict(VAR_SM,n.ahead = 15)

```

How does our model do predicting the long term?
```{r}
# preds$fcst$Sales ## useful reference - keep in 
# predictions long term 
preds = predict(VAR_SM,n.ahead = 15)
ase_var = mean((preds$fcst$Close[1:15] - mult_df_test$avgClose[1:15])^2)
ase_var
# rmse_var = sqrt(ase_var)
# rmse_var
print(paste0('The ASE for the VAR model, long term is: ', round(ase_var,4) ))
print(paste0('The square root of the ASE for the VAR model, long term is: ', round(sqrt(ase_var),4) ))
```


```{r}
# # plot forecast preds against actuals
plot(stock_weekly$avgClose[1:147], type='l',main = 'VAR Model LT Preds v Actuals',
      ylim = c(45,85))
lines(seq(143,147), predsST$fcst$Close[,1], col = 'blue')
```



MLP MODEL 
```{r}
trainDF = data.frame(FISCAL_WEEK_END = ts(mult_df_train$FISCAL_WEEK_END),
                     avgOpen = ts(mult_df_train$avgOpen),
                     avgHigh = ts(mult_df_train$avgHigh),
                     avgLow = ts(mult_df_train$avgLow),
                     avgVolume = ts(mult_df_train$avgVolume),
                     avgSales = ts(mult_df_train$avgSales),
                     avgTemp = ts(mult_df_train$avgTemp))
                     # avgFuel = ts(mult_df_train$avgFuel))
                     # avgCPI = ts(mult_df_train$avgCPI),
                     # avgUnemp = ts(mult_df_train$avgUnemp))

fit.mlp = mlp(ts(mult_df_train$avgClose), reps = 20, comb = 'mean', xreg=trainDF)
fit.mlp
plot(fit.mlp)
```

# MLP short term
```{r}
set.seed(17)
#  Sales
fit.mlp.sales = mlp(ts(mult_df_train$avgSales, frequency = 5), reps= 20, comb = "median")
plot(fit.mlp.sales)

fore.mlp.sales = forecast(fit.mlp.sales, h=5)
plot(fore.mlp.sales)

# temp
fit.mlp.temp = mlp(ts(mult_df_train$avgTemp, frequency = 5), reps= 20, comb = "median")
plot(fit.mlp.temp)

fore.mlp.temp = forecast(fit.mlp.temp, h=5)
plot(fore.mlp.temp)


df_fore = data.frame(week = ts(seq(1,114,1)),
                     temp = ts(c(mult_df_train$avgTemp, fore.mlp.temp$mean)),
                     sales = ts(c(mult_df_train$avgSales, fore.mlp.sales$mean)))
df_fore
multdf = data.frame(week = ts(mult_df_train$FISCAL_WEEK_END),
                    temp = ts(mult_df_train$avgTemp),
                    sales = ts(mult_df_train$avgSales))

fit.mlp = mlp(ts(mult_df_train$avgClose, frequency = 5), reps=20, comb = 'mean', xreg = multdf)
fit.mlp
plot(fit.mlp)

fore.mlp = forecast(fit.mlp, h=5, xreg = df_fore)
plot(fore.mlp)

ASE = mean((mult_df_test$avgClose[1:5] - fore.mlp$mean)^2)
ASE
```


```{r}
set.seed(17)
plot(seq(1,114,1), mult_df$avgClose[1:114], type='l', xlim = c(0,115), ylim = c(0,85))
lines(seq(109,113,1), fore.mlp$mean, type='l', col = 'red')
```

MLP rolling window RMSE: short term forecasts
```{r}
roll_mlp01ST = roll.win.rmse.wge(stock_train$avgClose, horizon = 5, phi = mod01est$phi,
                                  theta = mod01est$theta)

```

MLP Model: Short term forecast: Rolling window RMSE
```{r}
roll_mlp01ST_rmse = round(roll_mlp01ST$rwRMSE,4)
print(paste0('The rolling window RMSE for our MLP model, short term is: ', roll_mlp01ST_rmse))

```

# MLP long term
```{r}
set.seed(17)
#  Sales
# fit.mlp.sales = mlp(ts(mult_df_train$avgSales, frequency = 15), reps= 20, comb = "median")
# plot(fit.mlp.sales)

# use fit from above
fore.mlp.sales = forecast(fit.mlp.sales, h=15)
plot(fore.mlp.sales)

# temp
# fit.mlp.temp = mlp(ts(mult_df_train$avgTemp, frequency = 15), reps= 20, comb = "median")
# plot(fit.mlp.temp)

# use fit from above
fore.mlp.temp = forecast(fit.mlp.temp, h=15)
plot(fore.mlp.temp)


df_foreLT = data.frame(week = ts(seq(1,124,1)),
                     temp = ts(c(mult_df_train$avgTemp, fore.mlp.temp$mean)),
                     sales = ts(c(mult_df_train$avgSales, fore.mlp.sales$mean)))
df_foreLT

multdfLT = data.frame(week = ts(mult_df_train$FISCAL_WEEK_END),
                    temp = ts(mult_df_train$avgTemp),
                    sales = ts(mult_df_train$avgSales))

fit.mlpLT = mlp(ts(mult_df_train$avgClose, frequency = 15), reps=20, comb = 'mean', xreg = multdfLT)
fit.mlpLT
plot(fit.mlpLT)

fore.mlpLT = forecast(fit.mlp, h=15, xreg = df_foreLT)
plot(fore.mlpLT)

ASE_LT = mean((mult_df_test$avgClose[1:15] - fore.mlpLT$mean)^2)
ASE_LT
```

```{r}
set.seed(17)
plot(seq(1,124,1), mult_df$avgClose[1:124], type='l', xlim = c(0,125), ylim = c(0,120))
lines(seq(110,124,1), fore.mlpLT$mean, type='l', col = 'red')
```

MLP Model: Long term forecast: Rolling window RMSE
```{r}
roll_mlp01LT = roll.win.rmse.wge(stock_train$avgClose, horizon = 5, phi = mod01est$phi,
                                  theta = mod01est$theta)

```

```{r}
roll_mlp01LT_rmse = round(roll_mlp01LT$rwRMSE,4)
print(paste0('The rolling window RMSE for our MLP model, long term is: ', roll_mlp01LT_rmse))

```


# emsemble short term
```{r}
set.seed(17)
ensemble = (predsST$fcst$Close[,1] + fore.mlp$mean)/2
plot(seq(1,114,1), mult_df$avgClose[1:114], type='l', xlim = c(0,115), ylim = c(0,85))
lines(seq(109,113,1), ensemble, type='l', col = 'green')

ASE_ensST = mean((mult_df_test$avgClose[1:5] - ensemble)^2)
ASE_ensST
print(paste0('The ASE for the Ensemble model, short term is: ', round(ASE_ensST,4) ))
```


# emsemble long term
```{r}
set.seed(17)
ensembleLT = (preds$fcst$Close[,1] + fore.mlpLT$mean)/2
plot(seq(1,124,1), mult_df$avgClose[1:124], type='l', xlim = c(0,125), ylim = c(0,85))
lines(seq(110,124,1), ensembleLT, type='l', col = 'green')

ASE_ensLT = mean((mult_df_test$avgClose[1:15] - ensembleLT)^2)
ASE_ensLT
print(paste0('The ASE for the Ensemble model, long term is: ', round(ASE_ensLT,4) ))
```

FINAL CONCLUSION

RMSE Comparison of forecasts (on stock data, univariate & multivariate analysis).
```{r}
print(paste0('The rolling window RMSE for Model 01. AR(1), short term is: ', roll_mod01ST_rmse))
print(paste0('The rolling window RMSE for Model 01. AR(1), long term is: ', roll_mod01LT_rmse))
print(paste0('The rolling window RMSE for Model 02. ARIMA(1,2,1), short term is: ', roll_mod02ST_rmse))
print(paste0('The rolling window RMSE for Model 02. ARIMA(1,2,1), long term is: ', roll_mod01LT_rmse))
print(paste0('The rolling window RMSE for our MLP model, short term is: ', roll_mlp01ST_rmse))
print(paste0('The rolling window RMSE for our MLP model, long term is: ', roll_mlp01LT_rmse))
```

ASE Short term / Long term for VAR Model & Ensemble Model
```{r}
print(paste0('The ASE for the VAR model, short term is: ', round(ase_varST,4) ))
print(paste0('The square root of the ASE for the VAR model, short term is: ', round(sqrt(ase_varST),4) ))
print(paste0('The ASE for the VAR model, long term is: ', round(ase_var,4) ))
print(paste0('The square root of the ASE for the VAR model, long term is: ', round(sqrt(ase_var),4) ))
print(paste0('The ASE for the Ensemble model, short term is: ', round(ASE_ensST,4) ))
print(paste0('The square root of the ASE for the Ensemble model, short term is: ', round(sqrt(ASE_ensST),4) ))
print(paste0('The ASE for the Ensemble model, long term is: ', round(ASE_ensLT,4) ))
print(paste0('The square root of the ASE for the Ensemble model, long term is: ', round(sqrt(ASE_ensLT),4) ))
```


### IDEAS FOR FUTURE ANALYSIS 

Model 001. ARIMA(2,2,2) 

Difference the data again.
```{r}
# # pass in the differenced data
# mod02df1 = artrans.wge(mod02df, 1)
# plotts.sample.wge(mod02df1,arlimit=TRUE)
# aic5.wge(mod02df1)

```

```{r}
# mod02est = est.arma.wge(mod02df4,p=1, q = 1)
# 
# mean(mod02df1)
# mod02est$phi
# mod02est$theta
# mod02est$avar
```

Model 002. With moving average filter
Applying a moving average filter might be useful, as well.
```{r}
# # moving average filter (low-pass filter)
# library(forecast)
# ma_train01 = ma(stock_train$avgClose, order = 2)
```

Model 003. Multiple regression with correlated errors.

When the dependent and independent variables depend on time, it is common for the errors (residuals) to be auto-correlated.

Process Overview (MLE Approach)

Step 1. Regression analysis and model residuals.
Step 1a. Fit regression model.
Step 1b. Model ID on residuals.

Step 2. Use function ARIMA to perform MLE analysis which estimates the coefficient in the multiple regression while simultaneously modeling Z(t) as an AR(phi$p).

Use our new dataframe to model the data.
```{r}
# # Step 1a. Regression analysis.
# 
# # response = avgClose price 
# # explain variables = avgSales, avgFuel, avgCPI, avgUnemp
# 
# ksfit = lm(mult_df_train$avgClose ~ mult_df_train$avgSales + mult_df_train$avgFuel + mult_df_train$avgCPI + mult_df_train$avgUnemp)
# 
# # plot residuals
# # plot(ksfit$residuals, type='l')
# plotts.wge(ksfit$residuals)
# 
# # acf on residuals
# acf(ksfit$residuals)
# # residuals look in line with stationary behavior
# 
# parzen.wge(ksfit$residuals)

```

```{r}
# # model id on the residuals
# suppressWarnings(aic.wge(ksfit$residuals, p=0:17))
# # AIC picks p = 12, q = 2

```

```{r}
# AIC picks AR(1)

# next, store the parameter estimates in objects
# ksfit_params = suppressWarnings(aic.wge(ksfit$residuals, p=0:17))

```


```{r}
# fit ARIMA model using lagged variables with correlation ARMA(12,2)
# response = avgClose price 
# explain variables = avgSales, avgFuel, avgCPI, avgUnemp



```
